Alien civilizations: Good or bad?.  The spiritual factor


Indicators of the level of spirituality.



Since it is not possible to directly measure the spiritual factor, the
strategy would be to search for signs of the consequences of a moral or
spiritual nature in alien civilizations.


Lets call the indicator as "E-factor", standing for "spiritual factor".

In a first approach, civilizations can be spiritual based (E+) or with absence of spirituality (E-).



HOW DOES THE E-FACTOR AFFECTS THE SURVIVAL OF A CIVILIZATION?


Is it possible that an E+ civilization has more chances to survive
than an E- civilization?



CHARACTERISTICS


*** Civilizations with spiritual values (E+)

+ Seek for other civilizations to deepen its understanding of "divine plans"
+ Are tolerant to other emerging civilizations
+ Are tolerant to other ideas or organizations (?)
+ Are eager to share its vision about spirituality.

*** Civilizations without spiritual values (E-)

+ Pursue survival at all cost, even when it requires que conquer or enslave other species
+ Aggressive?


*** Characteristics independent of the E-factor

+ Use of natural resources
+ Development of science
+ Development of technology
+ Exploration of space

+ Is it possible for a E+ type civilization to have some of the E- type characteristics?   And viceversa?

+ Does the lifetime depend on the E-type?

+ Is it possible that an E+ civilization discovers a communication channel
that is unknown for E- civilizations?

WHY SEARCHING FOR ALIEN LIFE? (LISTENING)

+ Spiritual or religious reasons
+ Search for knowledge
+ Need of habitable locations or resources
+ monitor developing civilizations as a "galactic censorship"

WHY CALLING FOR ALIEN LIFE?

+ Spiritual or religious reasons
+ Search for knowledge
+ Need of habitable locations or resources
+ S.O.S. calls
+ Warnings for other civilizations about threats
+ Use messages as decoys to identify emerging civilizations


WHY NOT CALLING FOR ALIEN LIFE?

+ Fear of more powerfull and aggressive species looking for resources

COMMUNICATION CHANNELS

The emmision, reception and decodification are all necessary to complete qye
communication process.

+ What if all are listening but no one is talking?
+ Is it possible for an advanced civilization to avoid emmisions?


A STOCHASTIC PROCESS APPROACH

Hipothesis 1:  Aggressive spreading

Les supose that there are plenty of advanced civilizations in the
galaxy, some of them of E+ type and som of E- type.  If one type is
aggressive (E-) and the other is not (E+), would it mean (according to
"characteristics") that after some time E- types will prevail?

May be an aggressive civilization can not reach another distant
planet, although it knows its location and value.   Taking this into
account, the growth of aggressive species requires a fast travel
device.  Otherwise, its possible military power is not usefull.

If these civilizations discover a device to quickly travel to target
planets, the galaxy would be fully populated with this species and we
would have already made contact.   This evidence would then disfavour
this hypothesis?

Hipothesis 2:  The evil assumption

Given the total lack of knowledge about other species, a conservative
position would be to be a "quiet" planet, as long as it is possible,
just in case nearby civilizations are looking for planets with
valuable resources.   The fact that we have not detected any signs of
life in the universe outside Earth would favour this hypothesis but
can not be used to discard it or its opposite.

Hipothesis 3:  Spiritual independence

It is also possible that the survival of a civilization does not
depend at all on the development of a spiritual awareness.

It is also possible that the degree of hostility does not depend on
the E-factor, or that it is a consequence of a lack of undestanding or
sympathy towards less advanced species.  For example, humankind would
exploit resources of a solar system planet, even if it implies the
destruction of a less advanced "intelligent civilization" that we can
not undertand.  We would see them as "animals" and would wipe them out
if nececessary, albeit we can be considered as E+ type.


% Papers para citar y discutir:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

https://www.theguardian.com/science/2016/sep/23/stephen-hawking-warns-against-seeking-out-aliens-in-new-film




http://adsabs.harvard.edu/abs/2016arXiv161102586T







http://adsabs.harvard.edu/abs/1959Natur.184..844C

Searching for Interstellar Communications
by Cocconi, Giuseppe; Morrison, Philip


http://adsabs.harvard.edu/abs/2002PASP..114..416R

Optical Search for Extraterrestrial Intelligence: A Spectroscopic Search for Laser Emission from Nearby Stars
by Reines, Amy E.; Marcy, Geoffrey W.

http://adsabs.harvard.edu/abs/2001ARA%26A..39..511T

The Search for Extraterrestrial Intelligence (SETI)
by Tarter, Jill




 




%## GHZ

Morrison & Gowanlock
Extending Galactic Habitable Zone Modelling to Include the Emergence of Intelligent Life
https://ui.adsabs.harvard.edu/abs/2014ebi..conf..5.4M/abstract

Javaux & Dehant
Habitability: from stars to cells
https://ui.adsabs.harvard.edu/abs/2010A%26ARv..18..383J/abstract

Schwieterman et al.
A Limited Habitable Zone for Complex Life
https://ui.adsabs.harvard.edu/abs/2019ApJ...878...19S/abstract


%## Simulation parameters

%  D_max

%  tau_a

%  tau_s

Legassick 2015
The Age Distribution of Potential Intelligent Life in the Milky Way
https://ui.adsabs.harvard.edu/abs/2015arXiv150902832L/abstract




% hypotheses

W.I. Newman, C. Sagan 
Galactic civilizations: Population dynamics and interstellar diffusion
https://www.sciencedirect.com/science/article/pii/0019103581901354




%## SETI strategy: from causal contact to communication

% stars as amplifiers:

E.F. Borra & E. Trottier
Discovery of peculiar periodic spectral modulations in a small fraction of solar type stars
https://arxiv.org/abs/1610.03031

by Tamburini, Fabrizio; Licata, Ignazio
Can the periodic spectral modulations of the 236 SETI candidate Sloan Survey stars be due to Dark Matter effects?

Borra, Ermanno F.
Searching for Extraterrestrial Intelligence Signals in Astronomical Spectra, Including Existing Data
http://adsabs.harvard.edu/abs/2012AJ....144..181B


















